---
  title: "Working with Data in R"
---

```{r run_previous, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
pkgs <- c("tidyverse", "lubridate", "readxl", "highcharter", "knitr", "tidyquant", "timetk", "tibbletime", "quantmod", "PerformanceAnalytics", "scales")

sapply(pkgs, function(pkg) if(!require(pkg, character.only = TRUE)) install.packages(pkg, repos = "https://cloud.r-project.org/"))

sapply(pkgs, library, character.only = TRUE)
```

## Data Import

Naturally the first step is to import data into R. This is usually done using `read.csv` but if you have a .xls or .xlsx file, you will need to use `read_excel` from "readxl". I have a list of five stocks that we import into R using `read.csv`

To load our data into R, we use the read_csv() function from readr. Note that read_csv() would interpret our date column as a numeric but we can specify that our date column is in year-month-day format. We do that with `col_types = cols(Dates = col_date(format = "%Y-%m-%d"))`.

```{r read_csv, message=FALSE, results='hide', warning=FALSE}
filenames <- list.files(pattern = "*.csv")

prices <- filenames %>%
  map(~read_csv(.,
                col_types =
                  cols(Date =
                         col_date(format = "%Y/%m/%d")),
                na = c("", "#N/A", "NA"))) %>%
  bind_cols() %>%
  rename(date = `Date...1`) %>%
  select(date, everything(),-contains("...")) %>%
  tk_xts(date_var = date)

symbols <- substr(filenames, 1, 3)
```

Notice that the first line gets a list of filenames in the folder which have a .csv file extension. The filenames are then passed into a `map` call. This is essentially a loop under the hold, with the filenames being individually passed to the file argument of `read_csv()`. The `map` function returns a list of dataframes so `bind_cols` is used to make it one dataframe. Since this just combines the dataframes, the Dates column is duplicated. First, rename one of the duplicated Dates columns to date. Once that's done, select date first to put it as a first column, the everything else with the `everything()` helper function and drop the duplicated dates columns (since all duplicated column names will have "..." in their column names). 

The function tk(date_var = date) converts the data frame to an xts object. Once everything is all said and done, have one final look at the prices object. 

```{r head_code, results='hide'}
head(prices)
```



```{r prices_head, echo=FALSE}
# priceTable <- data.frame(index(head(prices)), head(prices))
priceTable <- data.frame(head(prices))
# colnames(priceTable)[1] <- ""
priceTable <- kable(priceTable, format = "html")
priceTable <- gsub("<table>", "<table class = kable>", priceTable)
priceTable
remove(priceTable)
```
