---
  title: "Building a Portfolio"
---
  
```{r run_previous, echo = FALSE, message=FALSE, results='hide', warning=FALSE}
pkgs <- c("tidyverse", "lubridate", "readxl", "highcharter", "knitr", "tidyquant", "timetk", "tibbletime", "quantmod", "PerformanceAnalytics", "scales")

sapply(pkgs, function(pkg) if(!require(pkg, character.only = TRUE)) install.packages(pkg, repos = "https://cloud.r-project.org/"))

sapply(pkgs, library, character.only = TRUE)

filenames <- list.files(pattern = "*.csv")

prices <- filenames %>%
  map(~read_csv(.,
                col_types =
                  cols(Date =
                         col_date(format = "%Y/%m/%d")),
                na = c("", "#N/A", "NA"))) %>%
  bind_cols() %>%
  rename(date = `Date...1`) %>%
  select(date, everything(),-contains("...")) %>%
  tk_xts(date_var = date)

symbols <- substr(filenames, 1, 3)

prices_monthly <- to.monthly(prices, indexAt = "lastof", OHLC = FALSE)

asset_returns <- Return.calculate(prices_monthly, method = "log") %>% na.omit()

asset_returns_long <- asset_returns %>%
  tk_tbl() %>%
  gather(asset, returns, -index) %>% 
  group_by(asset)
```

We spent a lot of time on the individual assets to make sure we had a good
grasp of our data building blocks.

Next, we collect individual returns into a portfolio, which is a weighted set of
asset returns. Accordingly, the first thing we need to do is assign a weight to
each asset. Our portfolio will have the following constituents:

+ ABG (Absa Group - Financials) weighted 25%
+ AGL (Anglo American Platinum - Resources) weighted 25%
+ GRT (Growthpoint - Property) weighted 20%
+ MRP (Mr Price - Retail) weighted 20%
+ NPN (Naspers - Media) weighted 10%

We need to create a weights vector that aligns with those allocations.

```{r portfolio_weights}
w <- c(0.25, .25, 0.20, .20, .10)
```

Before we use the weights in our calculations, have a quick sanity check in the
next code chunk to make sure the weights and assets align.

```{r sanity_check, results='hide'}
tibble(w, symbols)
```



```{r weights_table, echo=FALSE}
weights <- read.delim(text = 
"w symbols
<dbl> <chr>
0.250 SPY
0.250 EFA
0.200 IJS
0.200 EEM
0.100 AGG", sep = " ")

weights <- kable(weights, format = "html")
weights <- gsub("<table>", "<table class = kable id = 'weightsTable'>", weights)
weights
remove(weights)
```

Finally, make sure the weights sum to 100%. We can eyeball this with 5 assets,but with 50 assets it would be better to run the sanity check to catch any
errors as soon as possible

```{r unity}
tibble(w, symbols) %>% 
  summarise(total_weight = sum(w))
```

## Portfolio Returns in the xts world

We will use Return.portfolio() from PerformanceAnalytics, to calculate portfolio returns. The function requires two arguments for a portfolio, an xts object of returns and a vector
of weights. It is not necessary but we are also going to set rebalance_on =
"months" so we can confirm it matches our by-hand calculations above.
Remember, in the by-hand calculation, we set the portfolio weights as fixed,
meaning they never changed on a month-to-month basis. That is equivalent to
rebalancing every month. In practice, that would be quite rare. If we want a
more realistic scenario, we could choose annual rebalancing by changing the
argument to rebalance_on = "years".

```{r portfolio_returns, results=FALSE, message=FALSE, warning=FALSE}
portfolio_returns <- 
  Return.portfolio(asset_returns,
                   weights = w,
                   rebalance_on = "months") %>%
  `colnames<-`("returns")

head(portfolio_returns)
```



```{r porfolio_returns_table}
returns <- data.frame(portfolio_returns[1:5,])
colnames(returns)[1] <- ""
returns <- kable(returns, format = "html")
returns <- gsub("<table>", "<table class = kable>", returns)
returns
remove(returns)
```

## Portfolio Returns in the tidyverse

We begin our tidyverse work with our tidy data frame asset_returns_long.
Our first task is to add a weights column to the tibble using the mutate()
function. Each asset should be weighted according to the w vector. We use
case_when() to assign weights by asset, so that in the case when the asset
column is equal to asset 1 (ABG), we assign a weight of 0.25, or w[1], and so
on.

```{r asset_returns, results='hide'}
asset_returns_long %>%
  group_by(asset) %>%
  mutate(weights = case_when(asset == symbols[1] ~ w[1],
                             asset == symbols[2] ~ w[2],
                             asset == symbols[3] ~ w[3],
                             asset == symbols[4] ~ w[4],
                             asset == symbols[5] ~ w[5])) %>%
  head(6)
```



```{r tibble_table, echo=FALSE}
tibble_table <- read.delim(text = ',,,
# A tibble: 6 x 4,,,
# Groups:   asset [1],,,
  index,      asset,  returns, weights
  <date>,    <chr>,    <dbl>,   <dbl>
1 2000-02-29, ABG,   -0.0594,     0.25
2 2000-03-31, ABG,   -0.0248,     0.25
3 2000-04-30, ABG,    0.00418,    0.25
4 2000-05-31, ABG,    0.0606,     0.25
5 2000-06-30, ABG,    0.00781,    0.25
6 2000-07-31, ABG,   -0.0196,     0.25', sep = ",")

tibble_table <- kable(tibble_table, format = "html", col.names = NULL)
tibble_table <- gsub("<table>", "<table class = 'tibble', id = 'portTable'>", tibble_table)
tibble_table
remove(tibble_table)
```

Next, we need to implement the equation for portfolio returns. This task
is a bit tricky but serves as a nice way to use the group_by() function
with dates. We first add a new locum called weighted_returns that is the
product of each asset’s monthly return and its weight. Then, we group_by()
the date column because each of our weighted returns needs to be added
together for each date. Once we group by date, we can use summarise(total
= sum(weighted_returns)) to add up the monthly weighted returns.

```{r portfolio_returns_dplyr, results='hide'}
portfolio_returns_dplyr <- asset_returns_long %>%
  group_by(asset) %>%
  mutate(weights = case_when(asset == symbols[1] ~ w[1],
                             asset == symbols[2] ~ w[2],
                             asset == symbols[3] ~ w[3],
                             asset == symbols[4] ~ w[4],
                             asset == symbols[5] ~ w[5]),
         weighted_returns = returns * weights) %>%
  group_by(index) %>%
  summarise(returns = sum(weighted_returns))

head(portfolio_returns_dplyr, 6)
```



```{r port_Table_dplyr, echo=FALSE}
tibble_table <- read.delim(text = ',
,
# A tibble: 6 x 2,
  index,       returns
  <date>,        <dbl>
  2000-02-29, -0.0716 
  2000-03-31, -0.0232 
  2000-04-30, -0.0755 
  2000-05-31,  0.0528 
  2000-06-30,  0.00933
  2000-07-31,  0.00270', sep = ",")

tibble_table <- kable(tibble_table, format = "html", col.names = NULL)
tibble_table <- gsub("<table>", "<table class = 'tibble'>", tibble_table)
tibble_table
remove(tibble_table)
```

That piped workflow required some logical hoops but it useful to see how to
add those weights and then group by the date for finding total returns. Think
about how we would solve the puzzle of rebalancing weights not every month,
but every year.

## Portfolio Returns in the tidyquant world

In tidyquant, we start again with our long, tidy-formatted
asset_returns_long object, but convert to portfolio returns using
tq_portfolio().

The tq_portfolio function takes a tibble and then asks for an assets column
to group by, a returns column to find return data, and a weights vector.
It’s a wrapper for Return.portfolio() and thus also accepts the argument
rebalance_on = "months". Since we are rebalancing by months, we should
again get a portfolio returns object that matches our existing objects.

```{r tq_portfolio, results='hide'}
portfolio_returns_tq_rebalanced_monthly <-
  asset_returns_long %>%
  tq_portfolio(assets_col = asset,
               returns_col = returns,
               weights = w,
               col_rename = "returns",
               rebalance_on = "months")
```

Let’s take a quick look and compare how the tidy tibbles of portfolio returns compare to the xts objects of portfolio returns.

```{r compare, results = 'hide'}
portfolio_returns_dplyr %>%
  rename(tidyverse = returns) %>%
  mutate(tq = portfolio_returns_tq_rebalanced_monthly$returns,
         xts =coredata(portfolio_returns)) %>%
  mutate_if(is.numeric, list(~round(., 3))) %>%
  head(6)
```




```{r port_Table_comp, echo=FALSE}
tibble_table <- read.delim(text = ',,
,,
# A tibble: 6 x 4,,
  index,      tidyverse,     tq xts[,"returns"]
  <date>,         <dbl>,  <dbl>,           <dbl>
  2000-02-29,    -0.072, -0.072,          -0.072
  2000-03-31,    -0.023, -0.023,          -0.023
  2000-04-30,    -0.075, -0.075,          -0.075
  2000-05-31,     0.053,  0.053,           0.053
  2000-06-30,     0.009,  0.009,           0.009
  2000-07-31,     0.003,  0.003,           0.003', sep = ",")

tibble_table <- kable(tibble_table, format = "html", col.names = NULL)
tibble_table <- gsub("<table>", "<table class = 'tibble'>", tibble_table)
tibble_table
remove(tibble_table)
```

We have three objects of portfolio returns, calculated in three different ways, and
with the same results.

##  Visualizing Portfolio Returns in the xts world

Let us start with highcharter to visualize the xts formatted portfolio returns.
First, we set highchart(type = "stock") to get a nice
time series line. Then we add our returns column from
portfolio_returns. highcharter recognizes the
date index so we do not need to point to it.

```{r highcharts_port}
highchart(type = "stock") %>%
  hc_title(text = "Portfolio Monthly Returns") %>%
  hc_add_series(portfolio_returns$returns,
                name = "Rebalanced Monthly",
                color = "cornflowerblue") %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_navigator(enabled = FALSE) %>%
  hc_scrollbar(enabled = FALSE) %>%
  hc_legend(enabled = TRUE) %>%
  hc_exporting(enabled = TRUE)
```

The code results in a nice and simple line chart.

Let us see how well this portfolio did over time.

```{r highcharts}
c.port <- cumprod(1+portfolio_returns)

highchart(type = "stock") %>%
  hc_title(text = "Portfolio Monthly Returns") %>%
  hc_add_series(c.port$returns,
                name = "Rebalanced Monthly",
                color = "cornflowerblue") %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_navigator(enabled = FALSE) %>%
  hc_scrollbar(enabled = FALSE) %>%
  hc_legend(enabled = TRUE) %>%
  hc_exporting(enabled = TRUE)
```

Now let us compare that to the jse in the same time frame

```{r echo=FALSE, results='hide', warning=FALSE}
jse <- read_csv("data/jse.csv", col_types = cols(Date = col_date(format = "%Y/%m/%d")))
jse <- tk_xts(jse)
jse <- to.monthly(jse,  indexAt = "last", OHLC = FALSE)
jse <- Return.calculate(jse) %>% na.omit()
c.jse <- cumprod(1+jse)
```



```{r jse, warning=FALSE}
highchart(type = "stock") %>%
  hc_title(text = "Cumulative Returns") %>%
  hc_add_series(c.port[, 1], name = "portfolio") %>%
  hc_add_series(c.jse[, 1], name = "JSE") %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_navigator(enabled = FALSE) %>%
  hc_scrollbar(enabled = FALSE) %>%
  hc_exporting(enabled = TRUE) %>%
  hc_legend(enabled = TRUE)
```

The JSE outperforms our portfolio, looks like stock picking is hard, even with the benefit of hindsight. To be fair, our weighting scheme also has a lot to do with this. Overweighting Naspers might change this drastically.
